---
title: "Lab 16 - Neural Nets (Solution)"
author: NAME HERE
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
knitr::knit_meta(clean=T)
```

## 0 load the packages 
```{r}
#library() to load neuralnet, caret, and gains
library(neuralnet)
library(ggplot2)
library(lattice)
library(caret)
library(gains)

```

## 1 create a data frame
```{r}
# load the data 
df <- read.csv("EastWestAirlinesNN.csv")

# first six rows 
head(df)

# column names
names(df)
```

## 2 create two outcome dummies to represent output nodes  
```{r}
# create a variable named purchase

# create a variable named not_purchase

```

## 3 remove unnecessary variables
```{r}
# remove ï..ID column and update df 

# column names 

```

## 4 remove rows with missing values
```{r}


```

## 5 data partition 
```{r}
# set the seed 

# row indexes of the training set

# training set

# test set

```

## 6 normalize predictors to a scale of [0, 1] 
```{r}
# estimate the transformation 

# normalize the numerical predictors in the training set 

# normalize the numerical predictors in the test set 

```

## 7 fit a neural network on the training set 
```{r}
# set the random seed 

# fit a neural network that has a single hidden layer with 5 nodes 

# plot the network

# display weights

```

## 8 generate predicted probabilities for records in the test set  
```{r}
# make predictions using compute() 

# predicted probabilities of making purchases

```

## 9 generate predicted classes for records in the test set 
```{r}


```

## 10 create a confusion matrix for records in the test set
```{r}


```

## 11 create a gain table 
```{r}
# gain table 

# cumulative percentage of customers who make purchases 

# cumulative number of customers 

```

## 12 plot a lift chart
```{r}
# plot the cumulative number of customers who make purchases against the cumulative number of customers

# add a baseline curve 

```