---
title: "Lab 08 - Logistic Regression"
author: May Tsai
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## 0 load the package and suppress scientific notation 
```{r}
# load packages 
library(tidyverse)

# not display scientific notation
options(scipen=999)
```

## 1 create a data frame
```{r}
# load the data
ebay.df <- read.csv ("eBayAuctions.csv")
# first six rows
head(ebay.df)

# variable names 
names(ebay.df)
```

## 2 apply a function to each group 
```{r}
# mean of the binary outcome for each day of the week of auction close   
tapply(ebay.df$Competitive, ebay.df$endDay, mean)

# mean of the binary outcome for each auction category  

tapply(ebay.df$Competitive, ebay.df$Category, mean)


```

## 3 reduce levels in the day of week that the auction closed  
```{r}
# combine Sunday and Friday into a single category called Sun_Fri

Sun_Fri <- ebay.df$endDay[ebay.df$endDay == c("Sun","Fri")]
Sun_Fri



```

## 4 reduce levels in auction categories 
```{r}
# combine Business/Industrial and Computers into a single category called Computers
Computers <- ebay.df$Category[ebay.df$Category == c("Business/Industrial","Computer")]
Computers

# combine Antique/Art/Craft and Collectibles into a single category called collectibles

collectibles <- ebay.df$Category[ebay.df$Category == c("Antique/Art/Craft","Collectibles")]
collectibles




```

## 5 convert Duration to a categorical or factor variable 
```{r}
# convert Duration from a numeric variable to a categorical or factor variable  

ebay.df$Duration <- factor(ebay.df$Duration, levels = c("1", "3", "5", "7", "10"))
ebay.df$Duration

# variable types 

str(ebay.df)


```


## 6 data partition
```{r}
# set seed for reproducing the partition 
set.seed(1)

# takes a sample of row numbers for the training set 
train.index <- sample(c(1:dim(ebay.df)[1]), dim(ebay.df)[1]*0.6)  
head(train.index)

test.index <- sample(c(1:dim(ebay.df)[1]), dim(ebay.df)[1]*0.4)  
head(test.index)

# training set 
train.df <- ebay.df[train.index, ]
head(train.df)

# test set 
test.df <- ebay.df[test.index, ]
head(test.df)


```

## 7 fit a logistic regression model 
```{r,warning=FALSE}
# logistic regression using all the variables

logit.reg <- glm(Duration ~ (Category + currency + endDay), data = train.df, family = "binomial") 



# summary table 

summary(logit.reg)




```

## 8 generate predicted probabilities for records in the test set 
```{r}
# predicted probabilities 

logit.reg.pred <- predict(logit.reg, test.df, type = "response")



# first six values 

head(logit.reg.pred)

```

## 9 confusion matrix 
```{r}
confusionMatrix(as.factor(ifelse(logit.reg.pred > 0.5, 1, 0)),as.factor(test.df$Competitive.),positive = "1")

```

## 10 creating a gain table 
```{r}
# gain table 
library(gains)
head(test.df$Competitive)
head(logit.reg.pred)
gain <- gains(test.df$Competitive., logit.reg.pred, groups=10)
gain


# cumulative percentage of competitive auctions 
gain$cume.pct.of.total

# cumulative number of auctions 
gain$cume.obs
```

## 11 plot a lift chart
```{r}
# plot the cumulative number of competitive auctions against the cumulative number of auctions
plot(c(0,gain$cume.pct.of.total*sum(test.df$Competitive.))~c(0,gain$cume.obs), 
     xlab="cumulative number of competitive auctions", ylab="cumulative number of auctions", type="l")

# add a baseline curve 
lines(c(0,sum(test.df$Competitive.))~c(0, dim(test.df)[1]))
```