---
title: "Lab 03 - Data Wrangling"
author: May Tsai
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## 0 load the package 
```{r}
library(tidyverse)

```

## 1 create a tibble 

```{r}
# load the data 
firm<- read_csv('/home/rstudio/fin510/lab-03/firm.csv')
# first six rows 
head(firm)
# number of rows  
nrow(firm)
# number of columns 
ncol(firm)
# column names 
names(firm)
```

## 2 subset data 

### 2.1 one column 
```{r}
firm[[1]]

```

### 2.2 three columns 
```{r}
firm[,c(1,2,3)]
```

### 2.3 four rows and three columns
```{r}
firm[c(1:4),c("GVKEY","DATADATE","FYEAR")]

```

## 3 filter rows  

### 3.1 observations that belong to AAPL in 2015 
```{r}
filter(firm,FYEAR == '2015', TIC == 'AAPL')

```

### 3.2 observations that have the minimum NI value 
```{r}
min(firm$NI)

```

### 3.3 select columns of observations that belong to AAPL in 2015   
```{r}
firm[firm$FYEAR==2015 & firm$TIC=="AAPL",c("GVKEY","DATADATE","FYEAR","TIC","SIC","CONM" ,"AT","NI","DV")]

```

## 4 arrange rows 

### 4.1 ascending order of NI
```{r}
arrange(firm,NI)

```

### 4.2 descending order of NI
```{r}
arrange(firm,desc(NI))
```

### 4.3 ascending order of TIC and FYEAR, and descending order of NI
```{r}
arrange(firm,'TIC','FYEAR',desc(NI))

```

## 5 create new variables  

### 5.1 return on assets 
```{r}
#Create a new column named ROA in df which divides net income (NI) by total assets (AT).
firm$ROA<-firm$NI+firm$AT
head(firm)
```

### 5.2 net income bin numbers  
```{r}
# step1: identify thresholds of 4 bins with an equal number of observations 
NI_bin <- seq(min(firm$NI), max(firm$NI), (max(firm$NI) - min(firm$NI))/4)

# step2: remove the names of a named vector  
NI_bin <- unname(bins)
# step3: bin NI  
NI_bin
```


### 5.3 log of assets 
```{r}
#Create a new column named AT_LOG in df which computes the log value of total assets (AT) using mutate(). Use head() to print the first six rows of df. 
firm<-mutate(firm,AT_LOG=firm$NI+firm$AT) 
head(firm)
```


### 5.4 drop a column 
```{r}
#Remove return on assets (ROA) from df by assigning NULL to the column. Use head() to print the first six rows of df
firm$ROA<-NULL
head(firm)

```

