---
title: "Lab 04 - Data Exploration"
author: NAME HERE
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## 0 load the package 
```{r}
library(tidyverse)
```

## 1 create a tibble 

```{r}
# load the data 
firm<- read.csv('firm.csv')

# first six rows 
head(firm)

# column names
ncol(firm)

```

## 2 calculate summary statistics 

```{r}
# 2.1 average net income   

mean(firm$AT)

# 2.2 first decile of net income 
#Find a value of net income such that 10% net income values are less than it and 90% are greater than it using quantile().

quantile(firm$AT, c(0.9,0.1))

# 2.3 summary of net income
#Calculate the minimum, the first quartile, the median, the mean, the third quartile, and the maximum value of net income using summary()

summary(firm$AT)
 

```

## 3 compute counts and proportion of logical values    
```{r}
# 3.1 number of rows
# Count the number of observations that have dividends (DV) greater than 0 and net income (NI) greater than 10000 using sum()

sum(firm$DV > 0, firm$NI > 10000, na.rm=TRUE)


# 3.2 proportion of rows 
#Find the proportion of observations that have dividends (DV) greater than 0 and net income (NI) greater than 10000 using mean()


mean(c(firm$DV > 0, firm$NI > 10000), na.rm=TRUE)


```

## 4 summarize data with dpylr  

### 4.1 average and maximum net income 
```{r}
#Calculate the average value and the maximum value of net income, the number of observations, and the number of unique TIC values using summarize(). Specify the names of the summary statistics as NI_mean, NI_max, n_rows, and n_firms, respectively.

summarize(firm,NI_mean=mean(firm$NI),NI_max=max(firm$NI),n_rows=n(), n_firms=n_distinct(NI))

```

### 4.2 average and minimum net income by firm 
```{r}
#Group df by TIC and save the grouped data as df_by_TIC using group_by().Calculate the average value and the minimum value of net income by TIC, and the number of observations in each group using summarize(). Specify the names of the summary statistics as NI_mean, NI_min, and n_rows, respectively.

df_by_TIC <- 
  group_by(firm,TIC) %>%
  summarize(df_by_TIC,NI_mean=mean(df_by_TIC$NI),NI_min=min(df_by_TIC$NI),n_rows=n())

df_by_TIC

```

## 5 combine multiple operations with the pipe 

### 5.1 average net income by firm  
```{r}
#Group df by TIC, calculate average net income and save it as NI_mean, and arrange the result in a descending order of NI_mean.

df_by_TIC %>%
summarize(NI_mean=mean(df_by_TIC$NI))%>%
  arrange(desc(NI_mean))



```

### 5.2 most recent return on assets by firm 
```{r}
#Create a new variable named ROA, group df by TIC, order the result in a descending order of FYEAR, keep the first observation in each group, and name it as recent_ROA.Hint: return on assets (ROA) is defined as the ratio of net income (NI) by total assets (AT).


firm <- 
group_by(firm,TIC) %>%  
mutate(ROA = NI/AT) %>%
arrange(desc(firm,FYEAR)) %>%  
summarize(firm,recent_ROA=first(ROA))

firm

```

## 6 combine firm fundamentals with executives’ compensation	

### 6.1 load executive compensation data 
```{r}
#The file manager.csv contains executives’ compensation information regarding 7 firms during fiscal years of 2015 to 2018. We used this data set in lecture 3.
# load the data 
manager<- read.csv('manager.csv')

# first six rows 
head(manager)

# column names
ncol(manager)
```

### 6.2  frequency counts of firm
```{r}
#Return frequency counts of unique TIC values in df using table().Return frequency counts of unique TICKER values in manager using table(). Do these data sets cover same firms?

table(firm$TIC)
table(manager$TICKER)


```

### 6.3 frequency counts of year
```{r}
#Return frequency counts of unique FYEAR values in df using table().Return frequency counts of unique YEAR values in manager using table(). Do these data sets cover same fiscal years?

table(firm$FYEAR)
table(manager$YEAR)


```

### 6.4 inner join 
```{r}
# inner join 
merged <- inner_join(manager,firm,c("TICKER"="TIC","YEAR"="FYEAR"))
merged

# first six rows 
head(merged)


# column names 
names(merged)
```


### 6.5 average salary and return on assets during 2017 and 2018 
```{r}

merged %>%
  filter(YEAR==2017 & 2018) %>%
  mutate(ROA=NI/AT) %>%
  group_by(TICKER) %>%
  summarize(SALARY_mean=mean(SALARY),ROA_mean=mean(ROA),n_row=n()) %>%
  arrange(desc(ROA_mean))
  



```


